---
title: "lab_08"
author: "derek willis"
date: "8/26/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## You will need

* A Census API key

## Load libraries and establish settings

You'll need to load two packages for this: the tidyverse and tidycensus.

**Task** load these two packages

```{r}
# Turn off scientific notation
options(scipen=999)
library(tidyverse)
library(tidycensus)
```

## Setup Census API

You'll need your Census API Key:

**Task** set your API Key

```{r echo=FALSE}
#census_api_key("8aee6b62d6fc91f426b1707a4f0fdb75e7558ec8", install=TRUE)
```

## Load and view ACS variables

You'll also need to load the list of American Community Survey variables from the 2019 5-year estimates so you can identify the codes you need:

**Task** load the variables from the 2019 5-year ACS and save them to a dataframe

```{r}
# Write code here to load the ACS variables
acs5 <- load_variables(2019,"acs5", cache = TRUE)
```

## Load PPP data

You'll also need to load an RDS file of PPP loan applications that were removed from the original data by lenders in July.

**Task** read the RDS file and save it to a dataframe

```{r}
removed_ppp <- read_rds("data/lab_08.rds")
```

## Answer questions

**Q1** What is the Maryland jurisdiction with the lowest median age, according to the ACS 2019 5-year estimates? You will need to find the variable for median age first.
**A1** Baltimore City is the jurisdiction with the lowest median age, with a range from 35.6 to 35.2

```{r}
median_age <- get_acs(geography = "county",
                      variables = c("B01002_001"),
                      state = "MD",
                      year = 2019)
median_age %>% arrange(desc(estimate))
```

**Q2** Which Maryland counties have a median age of at least 50?
**A2** Talbot County and Worcester County

```{r}
#filtered from 45 just to make sure no county with a higher median age had a moe that would make this question inconclusive
median_age %>% filter(estimate > 45)
```

**Q3** Which Maryland counties have the highest percentage of 5-17 year-olds that speak Spanish at home? Start with the code you wrote for Q4 and add a second dataframe with the total population of 5-17 year-olds by county. For each dataframe, rename the `estimate` column to a meaningful name you've given the ACS variable, remove the `variable` columns in each using select and then join them based on GEOID and NAME. Then you can add a column to calculate the percentage of the total 5-17 population that speaks Spanish at home.
**A3** PG County, Montgomery County and Caroline County

```{r}
spanish_speaking_youth_pop <- get_acs(geography = "county",
                                      variables = "B16004_004",
                                      state="MD",
                                      year=2019)
youth_pop <- get_acs(geography = "county",
                     variables = "B16007_002",
                     state = "MD",
                     year = 2019)
colnames(spanish_speaking_youth_pop) <- c('GEOID', 'NAME', 'variable', 'population', 'moe')
colnames(youth_pop) <- c('GEOID', 'NAME', 'variable', 'population', 'moe')
spanish_speaking_youth_pop$variable <- NULL
youth_pop$variable <- NULL

spanish_speaking_youth_percentage <- spanish_speaking_youth_pop %>%
  inner_join(youth_pop, by=c("GEOID","NAME")) %>%
  rename(spanish_speaking_population=population.x, spanish_speaking_moe=moe.x, youth_population=population.y, youth_moe=moe.y) %>%
  mutate(min_spanish_speaking_pop=spanish_speaking_population-spanish_speaking_moe, max_spanish_speaking_pop=spanish_speaking_population+spanish_speaking_moe, min_youth_pop=youth_population-youth_moe, max_youth_pop=youth_population+youth_moe)%>%
  mutate(percentage=(spanish_speaking_population/youth_population)*100)
spanish_speaking_youth_percentage %>%
  select(NAME,percentage) %>%
  arrange(desc(percentage))
```

**Q4** Given the counties with the largest percentage of 5-17 year-olds that speak Spanish at home, what stands out to you about the data? How would you describe the top 5 counties? How would you describe Garrett County's numbers?
**A4** What stands out to me about the data is the relatively reliable moe for the top three counties and that this makes the percentage I found more reliable. That changes for the next two counties, which have highest moes relative to the estimated population, making the percentage unreliable. I woulld say that Garrett County's numbers are completely unusable because of the moe.

**Q5** Which 5-digit zip code had the most removed PPP applications, and what is the percentage of Black population in that zip code according to the ACS? To answer this question, you will need to creeate separate dataframes with ACS data on total population by zip code and Black population by zip code. In each dataframe, rename the `estimate` column and remove the `variable` column. You'll need to combine the two dataframes in order to calculate the percentage, and then you can combine it with the PPP data. Remember: the PPP data may not have only 5-digit zip codes!
**A5** The black population in the zip code 60620 is 95.7%.

```{r}
removed_ppp <- removed_ppp %>%
  mutate(zip5 = str_sub(zip, 1, 5))
removed_ppp %>%
  group_by(zip5) %>%
  summarise(
    count=n()
  ) %>%
  arrange(desc(count))
# 60620 had the most loans
chicago_pop <- get_acs(geography = "zcta",
                       variables = "B01003_001",
                       year = 2019)
chicago_black_pop <- get_acs(geography = "zcta",
                             variables = "B02001_003",
                             year=2019)
black_percentage <- chicago_black_pop %>%
  inner_join(chicago_pop, by=c("GEOID","NAME")) %>%
  rename(black_pop=estimate.x, black_moe=moe.x,total_pop=estimate.y, total_moe=moe.y) %>%
  mutate(percentage=(black_pop/total_pop)*100)
# I realized at the end of this question that I also could/should have joined the ppp data with the larger dataset. However, I just used the filter element in the dataset to find out information on each zip.
```

**Q6** How would you describe to an editor the top 10 zip codes with the most removed application in terms of Black population? Where are they located, and what do they have in common? Are there any zip codes that seem different, and if so why?
**A6** Many of the top 10 zip codes have large percentages of black population. Nine of the top 10 zip codes are in Chicago, while the zip code with the second-largest amount of removed loans is in Atlanta.

**Q7** What other columns from the PPP data could be useful in reporting on what happened to this removed loans?
**A7** The name column would be useful as a reporter could use it to find if there were any individuals who were applying for multiple loans. The lender column would also be useful. When filtering just for removed loans in the 60620 zip area, a number of lenders come up quite consistently. Many of these companies are also listed as sole proprietorships.
