---
title: "Reverse Engineering Project"
author: "Student names here"
date: "Date here"
output:
  html_document:
    theme: cerulean
    highlight: pygments
    toc: true
    toc_float:
      collapsed: true
      smooth_scroll: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction 

In this notebook, we are reverse engineering the story, "As police struggle to solve homicides, Baltimore residents see an ‘open season for killing’" (https://www.washingtonpost.com/investigations/as-police-struggle-to-solve-homicides-baltimore-residents-see-an-open-season-for-killing/2018/12/26/7ee561e4-fb24-11e8-8c9a-860ce2a8148f_story.html)

## Load libraries

Loading required libraries for this analysis.

```{r echo=FALSE, message=FALSE}
options(scipen=999)
library(tidyverse)
library(lubridate)
library(refinr)
library(tidycensus)
library(stringr)
# must put in your Census API key:
#census_api_key("8aee6b62d6fc91f426b1707a4f0fdb75e7558ec8", install=TRUE)
```

## Load and Cleaning Data

In this section, describe the source of the data, write a basic data dictionary for data you are working with, and discuss any caveats or issues you discovered working with this data. 

```{r}
# Load required data

# Path to data should be loaded from folder "data" i.e. read_csv("data/name_of_data.csv")

# Clean required data and prepare for analysis if needed. 

homicide_data <- read_csv("data/homicide-data.csv")

# This data was collected by Washington Post journalists totaling homicide info form 50 American cities. It was collected in different formats and different jurisdictions and was cross-referenced with FBI data to ensure accuracy. The data has already been standardized and cleaned by Washington Post reporters so the datasets were easy to manuveur. I double-checked the accuracy and consistency of spelling in OpenRefine.

```

## Sentences to Engineer
In this notebook, we are reverse engineering five sentences from the story.

### Sentence 1

* **Sentence text**: “As Baltimore has seen a stunning surge of violence, with nearly a killing each day for the past three years in a city of 600,000, homicide arrests have plummeted. City police made an arrest in 41 percent of homicides in 2014; last year, the rate was just 27 percent, a 14 percentage point drop.”
* **Analysis summary**: I was able to confirm the findings of this paragraph and data analysis by The Post. By separating the data from Baltimore into different years, it was then possible to summarise how many cases had been closed by arrest and divide that by the total number of homicides to find the arrest percentage.

```{r}
# Put code to reverse engineer sentence here
baltimore_homicides <- homicide_data %>%
  filter(city == "Baltimore")
# Here we use filter and str_detect to create 11 different datasets for each year of homicide data that we are provided.
o_seven_homicides <- baltimore_homicides %>%
  filter(str_detect(reported_date, "^2007"))
o_eight_homicides <- baltimore_homicides %>%
  filter(str_detect(reported_date, "^2008"))
o_nine_homicides <- baltimore_homicides %>%
  filter(str_detect(reported_date, "^2009"))
ten_homicides <- baltimore_homicides %>%
  filter(str_detect(reported_date, "^2010"))
eleven_homicides <- baltimore_homicides %>%
  filter(str_detect(reported_date, "^2011"))
twelve_homicides <- baltimore_homicides %>%
  filter(str_detect(reported_date, "^2012"))
thirteen_homicides <- baltimore_homicides %>%
  filter(str_detect(reported_date, "^2013"))
fourteen_homicides <- baltimore_homicides %>%
  filter(str_detect(reported_date, "^2014"))
fifteen_homicides <- baltimore_homicides %>%
  filter(str_detect(reported_date, "^2015"))
sixteen_homicides <- baltimore_homicides %>%
  filter(str_detect(reported_date, "^2016"))
seventeen_homicides <- baltimore_homicides %>%
  filter(str_detect(reported_date, "^2017"))
# Here we use group_by to isolate the disposition field and summarise the amount of cases that were "closed by arrest." We find that there were 86 cases where there were arrests in 2014 out of 211 total cases. That is an arrest percentage of 41 percent. 
fourteen_homicides %>%
  group_by(disposition) %>%
  summarise(arrested_amount = sum(disposition == "Closed by arrest")) %>%
  arrange(desc(arrested_amount))
86/211
# Here we run the same code for the 2017 set of data and find that 93 arrests were made from a total of 340 cases — a total of 27 percent.
seventeen_homicides %>%
  group_by(disposition) %>%
  summarise(arrested_amount = sum(disposition == "Closed by arrest")) %>%
  arrange(desc(arrested_amount))
93/340
# Display results of code below this codeblock

```

### Sentence 2

* **Sentence text**: "While homicide rates remain near historical lows in most American cities, Baltimore and Chicago are now both seeing murder tallies that rival the early 2000s."
* **Analysis summary**: I was able to confirm the finding that recent homicide numbers rivaled those of the early 2000s based on the total number of homicides seen each year. When comparing the results we can see a steady increase in homicides throughout each year. When looking at 2007 and 2017, there are significantly more homicides in the latter year, as 2007 saw 280 homicides and 2017 saw 340. 

```{r}
# Put code to reverse engineer sentence here
baltimore_homicides <- homicide_data %>%
  filter(city == "Baltimore")
#Here, we'll use filter and str_detect to create the same datasets from above for each year reported.
o_seven_homicides <- baltimore_homicides %>%
  filter(str_detect(reported_date, "^2007"))
o_eight_homicides <- baltimore_homicides %>%
  filter(str_detect(reported_date, "^2008"))
o_nine_homicides <- baltimore_homicides %>%
  filter(str_detect(reported_date, "^2009"))
ten_homicides <- baltimore_homicides %>%
  filter(str_detect(reported_date, "^2010"))
eleven_homicides <- baltimore_homicides %>%
  filter(str_detect(reported_date, "^2011"))
twelve_homicides <- baltimore_homicides %>%
  filter(str_detect(reported_date, "^2012"))
thirteen_homicides <- baltimore_homicides %>%
  filter(str_detect(reported_date, "^2013"))
fourteen_homicides <- baltimore_homicides %>%
  filter(str_detect(reported_date, "^2014"))
fifteen_homicides <- baltimore_homicides %>%
  filter(str_detect(reported_date, "^2015"))
sixteen_homicides <- baltimore_homicides %>%
  filter(str_detect(reported_date, "^2016"))
seventeen_homicides <- baltimore_homicides %>%
  filter(str_detect(reported_date, "^2017"))
#Now we can summarize each data set and see how many homicide cases there were for each year. If done right, we should see an increase. 
o_seven_homicides%>%
  summarise(o_seven_total=n())
o_eight_homicides%>%
  summarise(o_eight_total=n())
o_nine_homicides%>%
  summarise(o_nine_total=n())
ten_homicides%>%
  summarise(ten_total=n())
eleven_homicides%>%
  summarise(eleven_total=n())
twelve_homicides%>%
  summarise(twelve_total=n())
thirteen_homicides%>%
  summarise(thirteen_total=n())
fourteen_homicides%>%
  summarise(fourteen_total=n())
fifteen_homicides%>%
  summarise(fifteen_total=n())
sixteen_homicides%>%
  summarise(sixteen_total=n())
seventeen_homicides%>%
  summarise(seventeen_total=n())
# Display results of code below this codeblock

```

### Sentence 3

* **Sentence text**: As the murder rate soared, Mayor Catherine Pugh (D) fired Police Commissioner Kevin Davis in January and replaced him with Darryl De Sousa, who in turn resigned just four months later after facing federal charges of failing to file tax returns. De Sousa later pleaded guilty to the charges.
* **Analysis summary**: To confirm this finding, we need to find if the murder rate in the city was indeed increasing before and during Commissioner Kevin Davis' tenure in his position. We need to load population data for Baltimore and compare that to murder numbers in the city to find the murder rate. We were able to confirm the finding by the Post. In the year before Davis was commissioner, the homicide rate was around .034%. The year he took over, it jumped to .055% and hovered between that number and .052% for the next three years.

```{r}
# Put code to reverse engineer sentence here
# Davis was appointed commissioner in 2015 and was fired in Jan. 2019. Because of this, we will look at murder numbers from 2014-2017 to identify the murder rate. Before using the Census API, we need to load variables for the American Community Survey. We will use this data to load the population info for each year.
acs5 <- load_variables(2019,"acs5", cache = TRUE)
#This code will give us the population data for each of the years we are highlighting in Baltimore.
baltimore_population_14 <- get_acs(geography = "county",
                                   variables = "B01001_001",
                                   state = "MD",
                                   year = 2014)
baltimore_population_15 <- get_acs(geography = "county",
                                   variables = "B01001_001",
                                   state = "MD",
                                   year = 2015)
baltimore_population_16 <- get_acs(geography = "county",
                                   variables = "B01001_001",
                                   state = "MD",
                                   year = 2016)
baltimore_population_17 <- get_acs(geography = "county",
                                   variables = "B01001_001",
                                   state = "MD",
                                   year = 2017)
#This code takes out all the other counties in the datasets that we do not need. It keeps just Baltimore City.
baltimore_population_14 <- baltimore_population_14 %>%
  filter(NAME == "Baltimore city, Maryland")
baltimore_population_15 <- baltimore_population_15 %>%
  filter(NAME == "Baltimore city, Maryland")
baltimore_population_16 <- baltimore_population_16 %>%
  filter(NAME == "Baltimore city, Maryland")
baltimore_population_17 <- baltimore_population_17 %>%
  filter(NAME == "Baltimore city, Maryland")
#This code puts a city column in the population dataset with the value of Baltimore. This will be useful when we combine this dataset with the different years of murder data later.
baltimore_population_14 <- baltimore_population_14 %>%
  mutate(city = "Baltimore")
baltimore_population_15 <- baltimore_population_15 %>%
  mutate(city = "Baltimore")
baltimore_population_16 <- baltimore_population_16 %>%
  mutate(city = "Baltimore")
baltimore_population_17 <- baltimore_population_17 %>%
  mutate(city = "Baltimore")
#We are adding a column in the yearly homicide datasets that shows the total of homicides in that year. This will help us calculate the homicide rate with code when we combine the population and homicide data.
fourteen_homicides <- fourteen_homicides %>%
  mutate(total_murders = 211)
fifteen_homicides <- fifteen_homicides %>%
  mutate(total_murders = 342)
sixteen_homicides <- sixteen_homicides %>%
  mutate(total_murders = 320)
seventeen_homicides <- seventeen_homicides %>%
  mutate(total_murders = 340)
#We are joining the population data with the homicide data by the city column. Then we are adding a column called hom_rate which is total_murders divided by the estimated population for that year for each year we are looking at. The final colum in each dataset will be the murder rate for that year.
fourteen_homicides_pop <- fourteen_homicides %>%
  inner_join(baltimore_population_14, by=c("city")) %>%
  mutate(hom_rate = (total_murders/estimate)*100)
fifteen_homicides_pop <- fifteen_homicides %>%
  inner_join(baltimore_population_15, by=c("city")) %>%
  mutate(hom_rate = (total_murders/estimate)*100)
sixteen_homicides_pop <- sixteen_homicides %>%
  inner_join(baltimore_population_16, by=c("city")) %>%
  mutate(hom_rate = (total_murders/estimate)*100)
seventeen_homicides_pop <- seventeen_homicides %>%
  inner_join(baltimore_population_17, by=c("city")) %>%
  mutate(hom_rate = (total_murders/estimate)*100)
#We are writing code to simply view the homicide rate for each of these years. I have written as a comment the murder rate next to each year.
fourteen_homicides_pop %>% select(hom_rate) #.034%
fifteen_homicides_pop %>% select(hom_rate) #.055%
sixteen_homicides_pop %>% select(hom_rate) #.052%
seventeen_homicides_pop %>% select(hom_rate) #.055%
# Display results of code below this codeblock

```

### Sentence 4

* **Sentence text**: [Paste in sentence to engineer here]
* **Analysis summary**: [Write up two to three sentences describing the results of your analysis.  Were you able to confirm the finding? If not, why not?]

```{r}
# Put code to reverse engineer sentence here

# Display results of code below this codeblock

```

### Sentence 5

* **Sentence text**: [Paste in sentence to engineer here]
* **Analysis summary**: [Write up two to three sentences describing the results of your analysis.  Were you able to confirm the finding? If not, why not?]

```{r}
# Put code to reverse engineer sentence here

# Display results of code below this codeblock

```

-30-